---
title: "R Resources"
subtitle: "A collection of *free* resources for learning R and its packages. Contributions to this page are welcomed and encouraged! To submit a suggestion, please open a pull request on the repo. The only requirement to submissions is that they be free to access and use. Please ensure that any submission follows the same structure as others."
page-layout: full
title-block-banner: true
comments: false
---

```{r, echo = FALSE}
library(tidyverse)
library(gt)
library(gtExtras)
```

```{r, echo=FALSE}
gt_theme_site <- function(gt_object, ...) {
  
  table_id <- subset(gt_object[['_options']], parameter == 'table_id')$value[[1]]

  if (is.na(table_id)) {
    table_id <- gt::random_id()
    opt_position <- which("table_id" %in% gt_object[["_options"]][["parameter"]])[[1]]
    gt_object[["_options"]][["value"]][[opt_position]] <- table_id
  }

  gt_object %>%
    gt::tab_style(
      locations = gt::cells_body(
        columns = gt::everything()
      ),
      style = gt::cell_text(
        font = gt::google_font('Lora'),
        size = px(16)
      )
    ) %>%
    gt::tab_style(
      locations = gt::cells_column_labels(
        columns = gt::everything()
      ),
      style = gt::cell_text(
        font = gt::google_font('Lora'),
        weight = 650,
        size = px(16),
        transform = 'capitalize'
      )
    ) %>%
    gt::tab_style(
      locations = gt::cells_title('title'),
      style = gt::cell_text(
        font = gt::google_font('Lora'),
        weight = 650
      )
    ) %>%
    gt::tab_style(
      locations = gt::cells_title('subtitle'),
      style = gt::cell_text(
        font = gt::google_font('Lora'),
        weight = 500
      )
    ) %>%
    gt::tab_style(
      locations = gt::cells_source_notes(),
      style = gt::cell_text(
        font = gt::google_font('Lora'),
      )
    ) %>%
    gt::tab_options(
      heading.align = "left",
      column_labels.border.top.style = "none",
      table.border.top.style = "none",
      table_body.border.top.style = "none",
      table_body.border.bottom.color = "white",
      heading.border.bottom.style = "none",
      data_row.padding = px(3),
      column_labels.font.size = px(12),
      table.border.bottom.style = "none",
      source_notes.font.size = 10,
      source_notes.border.lr.style = "none",
      column_labels.border.bottom.style = "solid",
      column_labels.border.bottom.width = px(1),
      ...
    ) %>%
    gt::opt_css(
      paste0("#", table_id,
             " .gt_col_heading
             {
              padding-bottom: 3px;
             }",
             "#", table_id,
             " .gt_heading
             {
              padding-bottom: 0px;
              padding-top: 6px
            }",
             "#", table_id,
             " .gt_subtitle
             {
              padding-top: 2px;
              padding-bottom: 6px;
            }",
             "#", table_id,
             " .gt_sourcenote
             {
              line-height: 1.2
            }"),
      add = TRUE
    )
  
}
```

# ggplot2 (graphs)

<details>

<summary>`Tutorials/E-Books`</summary>

## E-Books

- [ggplot2: Elegant Graphics for Data Analysis](https://ggplot2-book.org) by **Hadley Wickham**, **Danielle Navarro**, and **Thomas Lin Pedersen**
  - A detailed introduction to `ggplot2` and its syntax/grammar. An excellent starting point to better understand the package as a beginner.
  
- [R Graphics Cookbook, 2nd edition](https://r-graphics.org) by **Winston Chang**
  - Includes more than 150 code snippets for basic `ggplot2` examples.

- [Data Visualization: A Practical Introduction](https://socviz.co) by **Kieran Healy**
  - Teaches the principles of effective data visualization, a brief introduction to `ggplot2`, and includes a nice focus on creating maps in R.

</details>
